---
title: "Sentiment Analysis: Wendy's vs. McDonald's"
author: "Levi Sayles"
date: "`r Sys.Date()`"
output: html_document
toc: true                 # Generates an automatic table of contents.
format:                   # Options related to formatting.
  html:                   # Options related to HTML output.
    code-tools: False      # Allow the code tools option showing in the output.
    embed-resources: TRUE # Embeds all components into a single HTML file. 
execute:                  # Options related to the execution of code chunks.
  warning: FALSE          # FALSE: Code chunk sarnings are hidden by default.
  message: FALSE          # FALSE: Code chunk messages are hidden by default.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidytext)
library(dplyr)
library(ggplot2)
library(readr)
library(lubridate)
```

## Introduction

In today’s competitive fast-food industry, understanding public sentiment is crucial for driving customer engagement and improving brand reputation. This report investigates the sentiment and emotions expressed in online discussions about two iconic fast-food chains: **Wendy's** and **McDonald's**. By leveraging sentiment analysis techniques and the NRC lexicon, we aim to uncover valuable insights into how these brands are perceived by their audiences.

### **Key Questions:**

1.  **Sentiment Comparison**: How do positive and negative sentiments differ between Wendy's and McDonald's?

2.  **Emotional Landscape**: What are the dominant emotions associated with each brand, and how do they vary?

3.  **Sentiment Trends**: How has sentiment for each brand evolved over time, and what factors might have influenced these trends?

The findings of this analysis will provide actionable insights to help these brands better understand their customers and optimize their marketing and public relations strategies.

## Data Collection

For this analysis, data was collected using **Google News RSS feeds**, capturing articles that specifically mention **"Wendy's"** and **"McDonald's"**. These articles provide unstructured textual data reflecting public discussions and opinions about each brand.

### **Processing Steps:**

1.  Articles were retrieved and stored in CSV files for analysis.

2.  The text data from titles and descriptions of the articles were combined to form a comprehensive text column for each brand.

3.  Sentiment and emotional context were analyzed using the **NRC lexicon** to categorize words into predefined sentiment and emotion classes.

This approach ensures that the analysis captures diverse and meaningful public discourse, offering an accurate view of brand perception over time.

```{r load-data}
#| include: FALSE 
# Load the datasets
wendys_news <- read_csv("https://myxavier-my.sharepoint.com/:x:/g/personal/saylesl_xavier_edu/EUoBrIj5OFpLsJBDpqXZrNwBcaMa0dJ9ecxlLNJhWi42Hg?download=1")
mcdonalds_news <- read_csv("https://myxavier-my.sharepoint.com/:x:/g/personal/saylesl_xavier_edu/EX4Xh8ewjE1JojZa464-BVIBS2zwX4DIt_juwczJX_IUOQ?download=1")

# Combine title and description into a single text column
wendys_news <- wendys_news %>%
  mutate(text = paste(title, description, sep = " "))

mcdonalds_news <- mcdonalds_news %>%
  mutate(text = paste(title, description, sep = " "))
```

## Question 1: Positive vs. Negative Sentiments

**Question:** Are discussions about Wendy's generally more positive or negative compared to McDonald's?

**Rationale:** This helps gauge overall public perception of each brand.

```{r sentiment-analysis}
#| echo: false
# Load NRC lexicon
nrc <- get_sentiments("nrc")

# Tokenize and join NRC lexicon
wendys_sentiment <- wendys_news %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc, by = "word", relationship = "many-to-many") %>%
  filter(sentiment %in% c("positive", "negative")) %>%
  count(sentiment, sort = TRUE)

mcdonalds_sentiment <- mcdonalds_news %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc, by = "word", relationship = "many-to-many") %>%
  filter(sentiment %in% c("positive", "negative")) %>%
  count(sentiment, sort = TRUE)

# Combine data for visualization
combined_sentiments <- rbind(
  data.frame(brand = "Wendy's", wendys_sentiment),
  data.frame(brand = "McDonald's", mcdonalds_sentiment)
)

# Visualize positive vs. negative sentiments
ggplot(combined_sentiments, aes(x = sentiment, y = n, fill = brand)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Positive vs Negative Sentiments",
       x = "Sentiment", y = "Count") +
  theme_minimal()
```

### Key Observations:

-   Wendy's has a higher proportion of positive sentiments compared to McDonald's.

-   Wendy's exhibits a slightly higher share of negative sentiments, highlighting areas for improvement.

### Implications:

-   Wendy's strong positive sentiment indicates a favorable public perception.

-   Wendy's should focus on reducing negative sentiment through improved customer engagement and public relations strategies.

## Question 2: Dominant Emotions

**Question:** What emotions dominate for Wendy's and McDonald's?

**Rationale:** Understanding emotional responses helps identify strengths and weaknesses in public perception.

```{r emotion-analysis}
#| echo: false 
# Count emotions
wendys_emotions <- wendys_news %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc, by = "word", relationship = "many-to-many") %>%
  count(sentiment, sort = TRUE)

mcdonalds_emotions <- mcdonalds_news %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc, by = "word", relationship = "many-to-many") %>%
  count(sentiment, sort = TRUE)

# Combine data for visualization
combined_emotions <- rbind(
  data.frame(brand = "Wendy's", wendys_emotions),
  data.frame(brand = "McDonald's", mcdonalds_emotions)
)

# Visualize emotions
ggplot(combined_emotions, aes(x = sentiment, y = n, fill = brand)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Dominant Emotions: Wendy's vs McDonald's",
       x = "Emotion", y = "Count") +
  theme_minimal()
```

### Key Observations:

**Positive Emotions:**

-   Wendy's has higher counts for "trust" than McDonald's, reflecting stronger public approval.

-   McDonald's shows significant "anticipation," suggesting customer excitement.

**Negative Emotions:**

-   Wendy's leads in "negative," "fear," and "sadness," indicating areas of concern or dissatisfaction.

-   McDonald's has lower counts for "disgust" compared to Wendy's.

**Implications:**

-   Wendy's can capitalize on its strong "trust" and "positive"sentiments.

-   Wendy's should investigate and address the root causes of "fear" and "sadness" to improve customer sentiment.

## Question 3: Sentiment Trends Over Time

**Question:** How has sentiment for Wendy's and McDonald's evolved over time?

**Rationale:** Tracking sentiment over time reveals trends and potential reactions to events.

```{r sentiment-trends}
#| echo: false
# Preprocess pubDate to remove the time zone
wendys_news <- wendys_news %>%
  mutate(pubDate = gsub(" GMT", "", pubDate))

mcdonalds_news <- mcdonalds_news %>%
  mutate(pubDate = gsub(" GMT", "", pubDate))

# Convert pubDate to Date type for Wendy's
wendys_news <- wendys_news %>%
  mutate(pubDate = as.Date(pubDate, format = "%a, %d %b %Y %H:%M:%S"))

# Convert pubDate to Date type for McDonald's
mcdonalds_news <- mcdonalds_news %>%
  mutate(pubDate = as.Date(pubDate, format = "%a, %d %b %Y %H:%M:%S"))

# Tokenize and group by date
wendys_trends <- wendys_news %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc, by = "word", relationship = "many-to-many") %>%
  filter(sentiment %in% c("positive", "negative")) %>%
  count(pubDate, sentiment)

mcdonalds_trends <- mcdonalds_news %>%
  unnest_tokens(word, text) %>%
  inner_join(nrc, by = "word", relationship = "many-to-many") %>%
  filter(sentiment %in% c("positive", "negative")) %>%
  count(pubDate, sentiment)

# Combine trends data
combined_trends <- rbind(
  data.frame(brand = "Wendy's", wendys_trends),
  data.frame(brand = "McDonald's", mcdonalds_trends)
)

# Visualize sentiment trends
ggplot(combined_trends, aes(x = pubDate, y = n, color = sentiment, linetype = brand)) +
  geom_line() +
  labs(title = "Sentiment Trends Over Time",
       x = "Date", y = "Sentiment Count") +
  theme_minimal()
```

### Key Observations:

-   Sentiment activity spikes around mid-November, indicating an event or campaign that drove discussions.

-   Wendy's shows stronger positive sentiment spikes compared to McDonald's.

-   Sentiment activity decreases toward the end of November, signaling reduced public discourse.

**Implications:**

-   Mid-November events for both brands significantly influenced public sentiment.

-   Wendy's may leverage its stronger positive sentiments in marketing campaigns.

## Conclusion

This analysis reveals significant differences in public sentiment and emotions between Wendy's and McDonald's:

-   Wendy's has stronger positive sentiments and higher trust and positive levels, giving it an advantage in public perception.

-   Wendy's also exhibits higher negative sentiments and emotions like fear and sadness, highlighting potential areas of improvement.

-   McDonald's exhibits higher anticipation, joy, and surprise, indicating that people are more excited overall about the changes brought to McDonald's than they are about Wendy's.

-   Both brands saw a significant spike in sentiment activity around mid-November, likely tied to specific events or campaigns.

These findings can guide marketing strategies and public relations efforts for both brands.
